Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.

### 2019-02-26 paper/snakemajoranas.tex
finish review with Anton halfway through sec V

%%██████████████████████████████████████████████████████████████████████████
%%██ Introduction
%%██████████████████████████████████████████████████████████████████████████
\section{Introduction}
\comment{Hybrid NS structures become topological and are useful for TQC.}
A hybrid structure containing a semiconductor with strong spin-orbit coupling coupled to a superconductor, inside a magnetic field becomes topological within the right parameter regime ($E_z^2>\mu^2+\Delta^2$ for a simple nanowire), and Majorana bound states appear on its edges.~\cite{lutchyn_majorana_2010,oreg_helical_2010}
Majorana bound states are a promising candidate to form the basis of a stable platform for topological quantum computing. ~\cite{alicea2012new,beenakker2013search}
The current experimental effort ~\cite{mourik_signatures_2012,das_zero-bias_2012,deng_anomalous_2012,churchill_superconductor-nanowire_2013,zhang2018quantized} focuses on creating pairs of Majorana bound states in hybrid normal-superconductor (NS) nanowire structures.

@@

\comment{We consider a two-dimensional system with zigzag and BdG Hamiltonian.}
We consider a Josephson junction (Fig.~\ref{fig:setup}) consisting of a 2D strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either saw-toothed zigzag as depicted [Fig.~\ref{fig:setup}(b)], or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG):
\begin{subequations}
\begin{align}
    H_\textrm{N} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu + \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \right] \tau_z
        + E_\text{z} \sigma_x, \\
    H_\textrm{SC} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z
        + \Delta \cos{\frac{\phi}{2}} \tau_x + \Delta \sin{\frac{\phi}{2}} \tau_y.
\end{align}
\label{eq:hamiltonian}
\end{subequations}
Here $H_\textrm{N}$ and $H_\textrm{SC}$ are the Hamiltonians of the semiconductor and superconductors, respectively.
The normal part has a linear Rashba spin-orbit coupling term with strength $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a coupling term $\Delta$, and the phases of the superconductors equal to $\pm\phi/2$.
Both the normal part and the superconductors have a kinetic term and chemical potential $\mu$.
The BdG Hamiltonian acts on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
We consider a zigzag pattern with a period $z_x$, a peak-to-peak amplitude $z_y$, and $W$ the width of the junction [see Fig.~\ref{fig:setup}(b)].
Later we relax this assumption and show that the exact shape is unimportant.

\comment{We discretize the Hamiltonian and simulate it with Kwant.}
We discretize our continuum Hamiltonian [Eq.~\eqref{eq:hamiltonian}] on a square grid and implement a tight-binding model using Kwant~\cite{groth_kwant:_2014}.
To preferentially sample important regions of parameter space, we use the Adaptive package.~\cite{adaptive}
The entire source code and the resulting raw data are available in Ref.~\onlinecite{data}.  % TODO: update the bib when the data is uploaded

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/wavefunctions}
\caption{Density of Majorana wave functions $\left| \psi_\textrm{M} \right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where lines parallel to a sinusoid defines the normal region, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for the straight system is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for the zigzag systems.
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across the zigzag geometries means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3.5 z_x=\SI{4550}{\nm}$, the remaining parameter values are written at the end of Sec.~\ref{sec:setup}.\label{fig:wave_functions}}
\end{figure}

@@

\comment{In a straight system, the Majoranas are very poorly localized because of small gap and high velocity}
We show the resulting Majorana wave function densities $\left| \psi_\textrm{M} \right|^2$ in different geometries in Fig.~\ref{fig:wave_functions} using the same Hamiltonian parameter values.
In the straight system [Fig.~\ref{fig:wave_functions}(a)], we see that the decay of the density is long compared to the system size.
The small topological gap combined with the high velocity result in a large Majorana size
\begin{equation}
\label{eq:xi_M}
\xi_\textrm{M}=\hbar \frac{v_\textrm{F}}{\Delta_\textrm{ind}},
\end{equation}
and therefore a minimal topological protection against perturbations.
The wave function extends to the center of the system, resulting in highly overlapping Majoranas and a Majorana coupling $E_\textrm{M}$ comparible to $E_\textrm{gap}$.

@@

\comment{We calculate the topological phase diagram using the gap size.}
In Fig.~\ref{fig:phasediagrams} we plot the energy gap as a function of magnetic field, chemical potential, $E_\textrm{gap}(B_x, \mu)$; and the superconducting phase difference $E_\textrm{gap}(B_x, \phi)$ for both a straight system [(c) and (e)] and a zigzag system [(d) and (f)].
Additionally, we plot the first $\SI{1300}{\nm}$ (one zigzag period) of the wave functions [(a) and (b)] at the most optimal point in parameter space for the straight system.
For the straight system we calculate $E_\textrm{gap}$ by performing a binary search in $E$ for the energy at which the propagating modes start to appear.~\cite{nijholt2015orbital}
Additionally, in Fig.~\ref{fig:phasediagrams}(c) we plot the phase boundaries that we find by solving a generalized eigenvalue problem as described in Ref.~\onlinecite{nijholt2015orbital}.
Due to the large size of the zigzag supercell, we are unable to use these full spectrum methods to compute these quantities.
Instead, we calculate $E_\textrm{gap}$ by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
By both observing the gap closings and knowing the topological phase for the non-zigzag system, we can infer the topology of the system with zigzag modulation.

>>>>

%%██████████████████████████████████████████████████████████████████████████
%%██ Introduction
%%██████████████████████████████████████████████████████████████████████████
\section{Introduction}
\comment{Hybrid NS structures become topological and are useful for TQC.}
A hybrid structure containing a semiconductor with strong spin-orbit coupling coupled to a superconductor, inside a magnetic field becomes topological within the right parameter regime ($E_z^2>\mu^2+\Delta^2$ for a simple nanowire), and Majorana bound states appear on its edges~\cite{lutchyn_majorana_2010,oreg_helical_2010}.
Majorana bound states are a promising candidate to form the basis of a stable platform for topological quantum computing. ~\cite{alicea2012new,beenakker2013search}
The current experimental effort ~\cite{mourik_signatures_2012,das_zero-bias_2012,deng_anomalous_2012,churchill_superconductor-nanowire_2013,zhang2018quantized} focuses on creating pairs of Majorana bound states in hybrid normal-superconductor (NS) nanowire structures.

@@

\comment{We consider a two-dimensional system with zigzag and BdG Hamiltonian.}
We consider a Josephson junction (Fig.~\ref{fig:setup}) consisting of a 2D strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either zigzag as depicted [Fig.~\ref{fig:setup}(b)], or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG):
\begin{subequations}
\begin{align}
    H_\textrm{N} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu + \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \right] \tau_z
        + E_\text{z} \sigma_x, \\
    H_\textrm{SC} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z
        + \Delta \cos{\frac{\phi}{2}} \tau_x + \Delta \sin{\frac{\phi}{2}} \tau_y.
\end{align}
\label{eq:hamiltonian}
\end{subequations}
Here $H_\textrm{N}$ and $H_\textrm{SC}$ are the Hamiltonians of the semiconductor and superconductors, respectively.
The normal part has a linear Rashba spin-orbit coupling term with strength $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a coupling term $\Delta$, and the phases of the superconductors equal to $\pm\phi/2$.
Both the normal part and the superconductors have a kinetic term and chemical potential $\mu$.
The BdG Hamiltonian acts on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
We consider a zigzag pattern with a period $z_x$, a peak-to-peak amplitude $z_y$, and $W$ the width of the junction [see Fig.~\ref{fig:setup}(b)].
Later we relax this assumption and show that the exact shape is unimportant.

\comment{We discretize the Hamiltonian and simulate it with Kwant.}
We discretize our continuum Hamiltonian [Eq.~\eqref{eq:hamiltonian}] on a square grid and implement a tight-binding model using Kwant~\cite{groth_kwant:_2014}.
To preferentially sample important regions of parameter space, we use the Adaptive package~\cite{adaptive}.
The entire source code and the resulting raw data are available in Ref.~\onlinecite{data}.  % TODO: update the bib when the data is uploaded

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{figures/wavefunctions}
\caption{Density of Majorana wave functions $\left| \psi_\textrm{M} \right|^2$ for sizes and geometries.
With (a) a straight system, (b) a zigzag system, (c) a system where lines parallel to a sinusoid defines the normal region, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for the straight system is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for the zigzag systems.
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across the zigzag geometries means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3.5 z_x=\SI{4550}{\nm}$, the remaining parameter values are written at the end of Sec.~\ref{sec:setup}.\label{fig:wave_functions}}
\end{figure}

@@

\comment{In a straight system, the Majoranas are very poorly localized because of small gap and high velocity}
We show the resulting Majorana wave function densities $\left| \psi_\textrm{M} \right|^2$ in different geometries in Fig.~\ref{fig:wave_functions} using the same Hamiltonian parameter values.
In the straight system [Fig.~\ref{fig:wave_functions}(a)], we see that the decay of the density is long compared to the system size.
The small topological gap combined with the high velocity result in a large Majorana size
\begin{equation}
\label{eq:xi_M}
\xi_\textrm{M}=\hbar \frac{v_\textrm{F}}{E_\textrm{gap}},
\end{equation}
and therefore a minimal topological protection against perturbations.
The wave function extends to the center of the system, resulting in highly overlapping Majoranas and a Majorana coupling $E_\textrm{M}$ comparible to $E_\textrm{gap}$.

@@

\comment{We calculate the topological phase diagram using the gap size.}
In Fig.~\ref{fig:phasediagrams} we compare the phase diagrams of the straight and the zigzag junction.
We plot $E_\textrm{gap}$ as a function of magnetic field, chemical potential, $E_\textrm{gap}(B_x, \mu)$; and the superconducting phase difference $E_\textrm{gap}(B_x, \phi)$ for both a straight system [(c) and (e)] and a zigzag system [(d) and (f)].
Additionally, we plot the first $\SI{1300}{\nm}$ (one zigzag period) of the wave functions [(a) and (b)] at the optimal point in parameter space for the straight system.
For the straight system we calculate $E_\textrm{gap}$ by performing a binary search in $E$ for the energy at which the propagating modes start to appear~\cite{nijholt2015orbital}.
Additionally, in Fig.~\ref{fig:phasediagrams}(c) we plot the phase boundaries obtained by solving a generalized eigenvalue problem~\cite{nijholt2015orbital}.
Due to the large size of the zigzag supercell, we are unable to use these full spectrum methods to compute these quantities.
Instead, we calculate $E_\textrm{gap}$ by finding the absolute minimum of the spectrum $E_\textrm{gap}=\min{|E(k)|}$.
By both observing the gap closings and knowing the topological phase for the non-zigzag system, we can infer the topology of the system with zigzag modulation.
